\sysfull

### 2018-08-03 paper/asplos19/abstract.tex
some cleaning up and partical commit

Energy harvesting, however, is unpredictable and can only power a device intermittently. Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve execution progress of a program. Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions. When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate. 
To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model. \sys progresses on a multi-tasks scale, when energy permits and preserves the computation progress on a sub-task scale if necessary. \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism. Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

>>>>

Energy harvesting, however, is unpredictable and can only power a device intermittently. Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program. Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions. When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate. 
To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model. \sys progresses on a multi-tasks scale when energy permits and preserves the computation progress on a sub-task scale if necessary. \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism. Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

### 2018-08-03 paper/asplos19/background.tex
some cleaning up and partical commit

Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
decompose their program into \textbf{tasks}, which is a region of code that can contain
arbitrary computation, sensing, and communication.  Task-based models keep
progress at the granularity of tasks by repeatedly trying to re-execute each
task on power failure until it successfully finishes, only then moving on to
the next task.  Since these models do not rely on taking an expensive
checkpoint, it is usually faster than the checkpoint-based
solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
task-based programming model, where the programmer explicitly expresses the
application as a sequence of tasks and the transitions between them.
%Task-based runtime systems ensure that repeated task executions are idempotent. 
%The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.  
%A programmer describes task control-flow as a {\em task graph}. Task flow happens at programmer demarcated transition points that may be conditional on program values. 
%Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.

Task-based models also suffer from data inconsistency when there is a WAR
dependence.  Prior systems tried to tackle the problem by a compiler-automated
redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
by statically creating a multiple copies of the problematic variable to ensure
that no task reads and writes the same copy~\cite{chain}.
%There are several run time strategies to ensuring task idempotence. One approach~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using private copies of those data, and commit the private copies on task completion. Another way is to statically create multiple versions of non-volatile data shared by tasks and ensure that no task reads and writes the same version~\cite{chain}. Regardless of the strategy, task-based systems execute statically-defined tasks atomically, completing in one or more attempt.

@@

\textbf{Frequent Non-volatile Memory Access.} 
Previous systems frequently access the non-volatile memory directly, which 
is a potential source of inefficiency.  Checkpointing systems copy all volatile
state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
every checkpoint.  Task-based models, which do not take checkpoints, back up a
subset of non-volatile data to maintain memory consistency using, e.g.,
redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
gets copied to and from the non-volatile memory in service of memory
consistency and progress preservation.  Ratchet~\cite{ratchet} and
Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
perform high frequency non-volatile memory accesses, into checkpointing
frequently.

%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.  
% We do not need to emphasize the following figure! (Sinan)
%\begin{figure}
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
%	\caption{Cost of \emph{write} operation}
%\end{subfigure}%
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
%	\caption{Cost of \emph{read} operation}
%\end{subfigure}
%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
%\end{figure}

% As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
% the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
% voltage drop on a capacitor powering the device across 1600 read and write
% accesses to each memory, randomizing accesses to avoid caching effects. Our
% data showed that a non-volatile memory access consumes around 1.5 times the
% energy of a volatile access. The data confirm that, because non-volatile memory
% accesses are more expensive than volatile memory accesses, prior systems that
% frequently access non-volatile memory may be artificially inefficient.

%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. This observation is also consistent with the TI specification~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}

% \sys is motivated by this observation: \sys 
% moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
% accesses, to decrease energy consumption and increase the time between power
% failures.

%The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.

%The above result reassures the intuitive observation that a decrease of checkpoint frequency (CP) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing. On the other hand, this increases task/checkpoint re-execution time. The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which however exposes a second problem.

%\begin{table}
%	\centering
%	\footnotesize
%	\begin{tabular}{|c|c|}
%		\hline
%		Platform name & Storage capacitor size \\
%		\hline \hline
%		%Moo~\cite{moo} & 0.1\,F \\
%		WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
%		%NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
%		NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
%		WISP~\cite{wisp5} & 47\,$\mu$F \\
%		{\em BioImpedance} sensor~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
%		{\em Ingestible} sensor~\cite{nadeau_naturebio_2017} & 220\,$\mu$F\\
%		\hline
%	\end{tabular} 
%	\caption{{\em Default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
%	%We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.
%	\label{table:capacitor}
%\end{table}

%Przemek placeholder
%\textbf{Task-based Models have Syntactic Limitations.} State of the art task-based programming models, Alpaca~\cite{alpaca}, has substantial limitations that prohibit writing generic programs. First, for Alpaca (and Chain~\cite{chain}), accesses to protected variables must be applied \emph{only} by using the variable's name, i.e. reference by variable address is not possible. This limitation implies that existing third-party libraries, such as MCU-tailored FFT library~\cite{}, cannot be implemented to support Alpaca or Chain (not to mention the impossibility of the development of an intermittent-safe \texttt{memcpy}). Furthermore, if a protected variable needs to be accessed in a code path that is not predictable by the compiler, for instance an Interrupt Service Routine, Alpaca cannot protect that access.

\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or
programmer-defined tasks can be both non-terminating and/or
inefficient.  If a task (or code between two checkpoints) consumes more than the fixed,
maximum energy that the device can buffer, then the task
will never be able to complete using buffered energy.  Such a task is
non-terminating, {\em prevents forward progress}, and makes the program
incorrect. 
%
If the task consumes far less energy than a device can buffer, the system
may operate {\em inefficiently}, saving the program state more often than needed.
%
Avoiding excessively costly, non-terminating, tasks and short, high-overhead,
tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}) are considered the problem of porting a program become much worse
because a large task on one device may be a relatively short task
on a device with a larger buffer.
%
\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which
merges multiple tasks to amortize the overhead when tasks are too small, and
{\em task downscaling}, which breaks up a task when the task is too large to complete. 
%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity. Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer. The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead. \sys's execution model based on {\em task coalescing} and {\em task downscaling}.  Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately. Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.

%\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware. This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency. \em{Fllexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology. \em{\sys champions an efficient, general, and flexible intermittent execution.}

>>>>

Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
decompose their program into \textbf{tasks}, which is a region of code that can contain
arbitrary computation, sensing, and communication.  Task-based models progress at the granularity of tasks. They re-execute each task interrupted by a power failure until it successfully finishes, only then moving on to
the next task. Since these models do not rely on taking an expensive
checkpoint, they are usually faster than the checkpoint-based
solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
task-based programming model, where the programmer explicitly expresses the
application as a sequence of tasks and the transitions between them.

Task-based models also suffer from data inconsistency when there is a WAR
dependence.  Prior systems tried to tackle the problem by a compiler-automated
redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
by statically creating multiple copies of the problematic variable to ensure
that no task reads and writes the same copy~\cite{chain}.
%There are several run time strategies to ensuring task idempotence. One approach~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using private copies of those data, and commit the private copies on task completion. Another way is to statically create multiple versions of non-volatile data shared by tasks and ensure that no task reads and writes the same version~\cite{chain}. Regardless of the strategy, task-based systems execute statically-defined tasks atomically, completing in one or more attempt.

@@

\textbf{Frequent Non-volatile Memory Access.} 
Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.  Checkpointing systems copy all volatile
state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
every checkpoint.  Task-based models, which do not take checkpoints, back up a
subset of non-volatile data to maintain memory consistency using, e.g.,
redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
gets copied to and from the non-volatile memory in service of memory
consistency and progress preservation.  Ratchet~\cite{ratchet} and
Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
perform high-frequency non-volatile memory accesses, into checkpointing
frequently.

\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or
programmer-defined tasks can be both non-terminating and/or
inefficient.  If a task (or code between two checkpoints) consumes more than the fixed,
maximum energy that the device can buffer, then the task
will never be able to complete using buffered energy.  Such a task is
non-terminating, {\em prevents forward progress}, and makes the program
incorrect. 
%
If the task consumes far less energy than a device can buffer, the system
may operate {\em inefficiently}, saving the program state more often than needed.
%
Avoiding excessively costly, non-terminating, tasks and short, high-overhead,
tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}) are considered the problem of porting a program become much worse
because a large task on one device may be a relatively short task
on a device with a larger buffer.
%
\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which
merges multiple tasks to amortize the overhead when tasks are too small, and
{\em task downscaling}, which breaks up a task when the task is too large to complete.

### 2018-08-03 paper/asplos19/systemdescription.tex
some cleaning up and partical commit

\sys is a new programming and execution model for intermittent computing on energy-harvesting devices. \sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable} and {\em efficient}. \sys accomplishes this goal with a constellation of a new programming model and run time software system support, that supports dynamically adaptive task-based execution. Figure~\ref{fig:system_overview} shows an overview of \sys.

\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by  designates a top-level set of functions as tasks, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
% In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.

% A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}. Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with automatically decomposed code would be identical.

The programmer compiles their task-based code, and links to \sys runtime,
producing a \sys-enabled binary. The runtime library relies on \sys's novel
{\em virtual memory manager} and its {\em adaptive task scheduler} to execute the binary 
efficiently.

\textbf{Adaptive Task Scheduler.} 
\sys's adaptive task scheduler decides when to coalesce consecutive tasks
together to amortize commit and transition costs, and when to downscale a task,
breaking it into multiple tasks, to avoid non-termination.

\sys's scheduler dynamically {\em coalesces} statically\hyp{}defined tasks to
avoid inessential overheads associated with completing tasks. By default, tasks
run in a sequence and each task commits its task-shared state as it completes.
The time and energy cost of a task's commit is unnecessary if the task and its
successor both complete without a power failure. The key insight that \sys
leverages is that the first task could have deferred its commit to the second
task. {\em Coalescing} tasks by deferring the first commit avoids the fixed
cost of the first commit (task transitioning overheads). Coalescing also
amortizes per-variable commit cost, committing each location accessed by both
tasks only after the second task, rather than once per task.  \sys makes
\emph{energy-aware} scheduling decisions, using a short history of recently
executed tasks to estimate available energy and guide future decisions. \sys
relies on {\em software only}, avoiding the need for additional circuitry to,
e.g., measure available energy.

%to set the adequate coalesced task
%size accordingly. This fundamental feature prevents \sys from blind coalesced
%task adaptation which may result in a repeated power failure problem. For
%example, If \sys enlarges the coalesced task size by $x$ static tasks on a
%coalesced task completion and reduces it by the same number of static tasks on
%a power reboot, then \sys may need to reboot, for instance, ten times to reduce
%a coalesced task size from ten to one, when $x=1$. Details of coalescing will
%be given in Section~\ref{sec:task_coalescing}.

>>>>

\sys is a new programming and execution model for intermittent computing on energy-harvesting devices. \sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable} and {\em efficient}. \sys accomplishes this goal with a constellation of a new programming model and runtime software system support, that supports dynamically adaptive task-based execution. Figure~\ref{fig:system_overview} shows an overview of \sys.

\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by encapsulating the code in a top-level set of functions, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
% In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.

% A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}. Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with the automatically decomposed code would be identical.

The programmer compiles their task-based code, and links to \sys runtime,
producing a \sys-enabled binary. The runtime library relies on \sys's novel
{\em virtual memory manager} and its {\em adaptive task scheduler} to adapt its execution with the energy conditions.

\textbf{Adaptive Task Scheduler.} 
\sys's adaptive task scheduler makes \emph{energy-aware} scheduling decisions to group tasks together to amortize commit and transition costs, or to downscale a task, breaking it into multiple tasks, to avoid non-termination. \sys's scheduler uses its recent execution history---no hardware dependency---as a metric to estimate energy availability.
